---
title: "Test gauge data"
author: "Galen Holt"
format:
  html:
    df-print: paged
---

```{r}
#| warning: false
#| message: false

# library(tidyverse)
library(sf)
library(reticulate) # Not strictly necessary but allows easier referencing of objects
```


I originally built this in Python ([scenario_creation.qmd](data_creation/scenario_creation.qmd)), but the only bit that *needs* to be in python is the gauge puller code. I'm adding this version as an example of how to mix R and python code chunks in similar situations where we only need a bit of code from the other language.

As long as the venv is in the base project directory, {reticulate} seems to find it. Otherwise, need to tell it where it is with `reticulate::use_virtualenv`.

I'm leaving the `=` alone rather than changing to `<-`.

# Creation of test data

We need to create test data for several scenarios for several gauges. This finds the gauges to pull based on location, gets the data, and modifies it into scenarios. The first case study is just the Macquarie, but I think I'd also like some additional catchments so the spatial aggregation demonstrations are more interesting. Is there any reason not to do that? Not really.

The shapefiles used to see what we're doing and do the selecting were produced with `data_creation/spatial_data_creation.qmd`.

I output clean test data to `data`, following R package data conventions. But set a 'datadir', so we can change that easily to somewhere else if desired.

First, do we actually want to rebuild the `data/` directory, or just use this to investigate and document what happened?

```{r}
REBUILD_DATA = FALSE
```

Set the data dir to make that easy to change.

```{r}
datarawdir = 'data-raw'
dataoutdir = 'data'
```

## Read in gauges

Use the bom gauges, since this is what were contained in the EWR tool.

```{r}
gaugegeo = read_sf(file.path(dataoutdir, 'bom_gauges.shp'))
```

## Read in polygons

I'll read in a few options, and then choose- basin, resource plan areas, sdl areas (which is where David got the Macquarie polygon), and catchments. I have cut out the testing of David's single polygon- it is the Macquarie-Castlereagh in the sdls.

```{r}
basin = read_sf(file.path(dataoutdir, 'basin.shp'))
rps = read_sf(file.path(dataoutdir, 'resource_plan.shp'))
sdl = read_sf(file.path(dataoutdir, 'sdl_units.shp'))
ltv = read_sf(file.path(dataoutdir, 'cewo_valleys.shp'))
```

crs all match from the creation.

### Plot the polygons and data checks

Basin

```{r}
ggplot(basin) + geom_sf(fill = 'powderblue')
```

Resource plan areas

```{r}
ggplot(rps) + geom_sf(aes(fill = SWWRPANAME), show.legend = FALSE) +
  geom_sf_label(aes(label = SWWRPANAME), size = 3, label.padding = unit(0.1, 'lines')) + 
  colorspace::scale_fill_discrete_qualitative(palette = 'Set2')
```

These have 'SW' codes

```{r}
rps
```

SDL plan areas

```{r}
ggplot(sdl) + geom_sf(aes(fill = SWSDLName), show.legend = FALSE) +
  geom_sf_label(aes(label = SWSDLName), size = 3, label.padding = unit(0.1, 'lines')) + 
  colorspace::scale_fill_discrete_qualitative(palette = 'Set2')
```

These have 'SS' codes.

```{r}
sdl
```

Catchments

```{r}
ggplot(ltv) + geom_sf(aes(fill = ValleyName), show.legend = FALSE) +
  geom_sf_label(aes(label = ValleyName), size = 3, label.padding = unit(0.1, 'lines')) + 
  colorspace::scale_fill_discrete_qualitative(palette = 'Set2')
```

```{r}
ltv
```

## Cut to demo polygons

The Macquarie shapefile from David is the Macquarie-Castlereagh from `rps`. I want a few catchments to play with for demos, so let's use the Macquarie, Castlereagh, Namoi, Lachlan. That might be a LOT of gauges, though.

```{r}
catch_demo = rps %>% 
  dplyr::filter(SWWRPANAME %in% c('Macquarie-Castlereagh', 'Lachlan', 'Namoi'))
```

```{r}
(ggplot(catch_demo) + geom_sf(aes(fill = SWWRPANAME)) +
scale_fill_brewer(type = 'qual', palette = 8))
```

## Get relevant gauges

### Cut to the polygons

Cut the gaugegeo from the whole country to just those four catchments

```{r}
demo_gauges = st_intersection(gaugegeo, catch_demo)
```

How many are we talking?

```{r}
demo_gauges %>% nrow()
```

295 rows is a lot, but unlikely they'll all be in the EWR.

### Extract their names

To feed to the gauge extractor, we need their gauge numbers.

```{r}
gaugenums = demo_gauges$gauge
```

## Get the values

We can feed lists to `gg.gauge_pull`, so can feed it that way. We might *want* to loop for parallelising extraction or modifications, but the real scenarios won't be made this way anyway, so not worth it here.

What time span do we want? 10 years to start

```{r}
starttime = lubridate::ymd(20100101)
endtime = lubridate::ymd(20191231)
```

How many are actually in the EWR tool? I could go get the table myself, but the EWR tool has a function, so use that. *This is where we have to use python to access the EWR table*. 

*Access the python objects with `py$objname`. 

```{python}
from py_ewr.data_inputs import get_EWR_table
ewrs, badewrs = get_EWR_table()
```

What are those gauges, and which are in both the ewr and the desired catchments?

```{r}
ewrgauges = py$ewrs$Gauge
ewr_demo_gauges = gaugenums[gaugenums %in% ewrgauges]
length(ewr_demo_gauges)
```

47 seems ok. Let's go with that.

### Get all the gauge data

Now we have a list of gauges, go actually get their hydrographs. Takes a while, be patient.

*Again, need python here.* *Pass R objects to python as* `r.`

*It takes -forever- to do a type translation on the* `DATETIME` *column*. So change it to something simple while still in python, and change it back in R next.

```{python gaugepulling}
#| message: false
import mdba_gauge_getter as gg
demo_levs = gg.gauge_pull(r.ewr_demo_gauges, start_time_user = r.starttime, end_time_user = r.endtime)
demo_ids = demo_levs.SITEID.unique()
len(demo_ids)

# I think this will work, the above is running
demo_levs['Date'] = demo_levs['DATETIME'].astype(str)
```

I guess that's not terrible. 168k rows is fine in general, but likely overkill for testing. I could cut it to fewer gauges and less time, probably, but it will be good to have reasonable time periods for testing time windowing if this doesn't eat too much time elsewhere.

For some reason `demo_levs['VALUE']` is an object and not numeric. And `'DATETIME'` needs to be named `Date` for the EWR tool to read it.

*I'm copying the py obj to R for further manipulation.* A bit dumb from a memory point of view, but it's not big.

Transforming the python dates to R dates is horribly slow. MUCH faster to change to string and back again. 

```{r datatransfer}
demo_levs <- py$demo_levs

demo_levs$VALUE = as.numeric(demo_levs$VALUE)

# # In python, we just need to change the name of the date column. Here, we need to change the python datetime.date objects to R dates
# 
# # Would this be faster if we made them numeric in pandas, sent them over that way, and then re-dated them? Almost certainly
# demodates <- demo_levs$DATETIME
# 
# # Really slow... I'm sure could speed up with furrr or something.
# rdates <- purrr::map(demodates, py_to_r) %>% 
#   tibble(.name_repair = ~'Date') %>%  
#   unnest(cols = Date)
# 
# 
# demo_levs <- bind_cols(rdates, demo_levs)
demo_levs <- dplyr::select(demo_levs, -DATETIME) %>% 
  dplyr::mutate(Date = lubridate::ymd(Date))
```


### Map the gauges

```{r}
# gaugegeo <- gaugegeo %>% 
#   dplyr::rename(gauge = `gauge number`)
demo_geo = gaugegeo %>% dplyr::filter(gauge %in% py$demo_ids)
```

Looks reasonable. Probably overkill for testing, but can do a cut down version too.

```{r}
(ggplot() + 
geom_sf(data = basin, fill = 'lightsteelblue') +
geom_sf(data = catch_demo, mapping = aes(fill = SWWRPANAME)) +
geom_sf(data = demo_geo, color = 'black') +
scale_fill_brewer(type = 'qual', palette = 8))
```

## Make test scenarios

### Demo scenarios

For now, the test scenarios are just multiplying by 4 or 0.25 to give something to work with. This section could easily be modified for other simple scenarios.

```{r}
down4 = demo_levs
up4 = demo_levs

down4$VALUE = down4$VALUE * 0.25
up4$VALUE = up4$VALUE * 4
```

### Make the data look like IQQM

I read this in with the EWR tool's code, but need to be able to read into the EWR tool through the `scenario_handling`. That's sort of convoluted, and could be fixed by splitting up `process_scenarios` into a read-in bit and a bit that calls `evaluate_EWRs.calc_sorter`. The end of `process_scenarios` starting with `gauge_results = {}` is the same as `process_gauges` and so could be shared.

For now though, just tell EWR it's IQQM so I can use `scenario_handling`. As long as I have a date column `Date` and other columns with gauge numbers for names, I can get the EWR tool to work by telling it it's IQQM. It can have multiple gauges, with each having its own column. I had originally set up for single files per scenario-gauge combination, but let's use multi-cols for now. The read-in should work either way.

### Save the output

Not being too fussed about structure, since the structure is going to change once we have actual scenarios.

I will still do the dir/file structure even with one file as it allows easier changes and more flexibility.

```{r}
scenenames = c('base', 'up4', 'down4')

# the full scenarios
for (x in scenenames) {
      scenedir = file.path(dataoutdir, 'testscenarios', x)
    if (!dir.exists(scenedir)) {
      dir.create(scenedir, recursive = TRUE)
    }
        
}


# a minimal set
for (x in scenenames) {
      scenedirS = file.path(dataoutdir, 'testsmall', x)
    if (!dir.exists(scenedirS)) {
      dir.create(scenedirS, recursive = TRUE)
    }
}

```

Create clean dataframes to save.

```{r}
base <- demo_levs %>% 
  dplyr::select(Date, VALUE, SITEID) %>% 
  tidyr::pivot_wider(id_cols = Date, names_from = SITEID, values_from = VALUE)

up4 <- up4 %>% 
  dplyr::select(Date, VALUE, SITEID) %>% 
  tidyr::pivot_wider(id_cols = Date, names_from = SITEID, values_from = VALUE)

down4 <- down4 %>% 
  dplyr::select(Date, VALUE, SITEID) %>% 
  tidyr::pivot_wider(id_cols = Date, names_from = SITEID, values_from = VALUE) 
```

Save. could do this above easily enough, but getting lots of dots and hard to read.

```{r}
if (REBUILD_DATA) {
      write_csv(base, file.path(dataoutdir, 'testscenarios', 'base', 'base.csv'))
      write_csv(up4, file.path(dataoutdir, 'testscenarios', 'up4', 'up4.csv'))
      write_csv(down4, file.path(dataoutdir, 'testscenarios', 'down4', 'down4.csv'))
}

```

### Smaller set

I also want a very limited set of test data that runs faster through the EWR tool and the rest of the toolkit. It still needs to have \> 1 gauge in \> 1 catchment though. Let's do three gauges in two catchments, 5 years.

Which gauges?

```{r}
mac = rps %>% 
  dplyr::filter(SWWRPANAME %in% c('Macquarie-Castlereagh'))
lach = rps %>% 
  dplyr::filter(SWWRPANAME %in% 'Lachlan')

macgauge = st_intersection(demo_geo, mac)
lachgauge = st_intersection(demo_geo, lach)
```

I'll just choose three of each

```{r}
macsub = macgauge %>% 
  dplyr::filter(gauge %in% c('421004', '421001', '421011'))
lachsub = lachgauge %>% 
  dplyr::filter(gauge %in% c('412005', '412002', '412038'))

minigauge = bind_rows(macsub, lachsub)

minigauges = unique(minigauge$gauge)
```

```{r}
(ggplot() + 
geom_sf(data = basin, fill = 'lightsteelblue') +
geom_sf(data = catch_demo, mapping = aes(fill = SWWRPANAME)) +
geom_sf(data = minigauge, color = 'black') +
scale_fill_brewer(type = 'qual', palette = 8))
```

```{r}
if (REBUILD_DATA) {
  start15 = lubridate::ymd(20150101)

    base %>% 
      dplyr::filter(Date > start15) %>% 
      dplyr::select(Date, all_of(minigauges)) %>% 
      write_csv(file.path(dataoutdir, 'testsmall', 'base', 'base.csv'))
    
    down4 %>% 
      dplyr::filter(Date > start15) %>% 
      dplyr::select(Date, all_of(minigauges)) %>%  
      write_csv(file.path(dataoutdir, 'testsmall', 'down4', 'down4.csv'))
    
    up4 %>% 
      dplyr::filter(Date > start15) %>% 
      dplyr::select(Date, all_of(minigauges)) %>%  
      write_csv(file.path(dataoutdir, 'testsmall', 'up4', 'up4.csv'))
}
    

```


